---
title: "Lab 1 - Data visualization"
format: html
editor: visual
author: "Derek Dubrule"
---

# Questions

## Part 1

```{r}
#| label: load packages

library(tidyverse)
```

### Question 1

```{r}
# Plot Number 1
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100"
)

# Plot Number 2
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 1000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 1000"
)

# Plot Number 3
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 10000"
)

# Plot Number 4
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100000"
)

```

### Question 2

```{r}
# Boxplot
ggplot(midwest, aes(x = popdensity)) +
  geom_boxplot() +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties"
)

```

The Distribution of population density among the Midwestern counties has the majority of the values on the left so it is right skewed. Having the majority of the counties on the left means that most have a relatively low population density. But there are several data points that have higher population densities. This pattern is consistent with what was observed in the histograms from Question 1.

The Boxplot shows several high population density outliers. Using the View(midwest) command in the Console, I was able to examine the data more closely and pick out which county was the furthest right. The furthest outlier in the data set is Cook Country, Illinois which as a population density of 88018.3966 which is significantly higher compared to most midwestern counties. Doing a quick Google search it was discovered that this county contains the city of Chicago so this makes sense as this is a major urban city with a large amount of people living in close conditions.

### Question 3

```{r}
# Scatterplot
ggplot(midwest, 
       aes(
         x = percollege,
         y = percbelowpoverty,
         color = state,
         shape = state)
       ) +
  geom_point() +
  labs(
    x = "Percentage of population with a college degree",
    y = "Percentage of population below poverty",
    title = "Percentage below poverty vs college education in Midwestern counties",
    color = "State",
    shape = "State"
  )
  
```

The plot that was generated shows that most counties located in the bottom left portion, indicating that many Midwestern counties have low percentages of both college educated residents and people living below the poverty line. Despite this clustering in the bottom left, there is still a slight negative relations between people with a college degree and the percentage of people below poverty. This means that counties with higher levels of college education generally tend to have lower poverty rates.

There a few outliers in this data. The first is a county in WI, which appears in the top left portion of the plot, indicating a high poverty rate and a low percentage of the population with a college education. Another outlier is a county in IL. I would assume this is the Cook County as we discussed in the previous question which contains Chicago. This point is located the furthest top right in respect to the other points. This position on the graph shows a higher level of education then most countys but also a higher level of poverty levels then most counties. This could be due to the fact that there is a wide range of education and wages for the people who live in a large metropolitan area.

### Question 4

### Question 5

### Question 6

### Question 7

## Part 2

**Enough about the Midwest!**

### Question 8

### Question 9
